---
title: "Mini Project 1"
author: "Edvin Stranberg"
date: "Jan 29, 2025"
---

```{r}
library(tidyverse)
```

```{r}
## create population graphs

norm_df <- tibble(x = seq(3, 17, length.out = 1000),
                  dens = dnorm(x, mean = 10, sd = 2),
                  pop = "normal(10, 4)")
unif_df <- tibble(x = seq(7, 13, length.out = 1000),
                  dens = dunif(x, 7, 13),
                  pop = "uniform(7, 13)")
exp_df <- tibble(x = seq(0, 10, length.out = 1000),
                 dens = dexp(x, 0.5),
                 pop = "exp(0.5)")
beta_df <- tibble(x = seq(0, 1, length.out = 1000),
                  dens = dbeta(x, 8, 2),
                  pop = "beta(8, 2)")

pop_plot <- bind_rows(norm_df, unif_df, exp_df, beta_df) |>
  mutate(pop = fct_relevel(pop, c("normal(10, 4)", "uniform(7, 13)",
                                  "exp(0.5)", "beta(8, 2)")))

ggplot(data = pop_plot, aes(x = x, y = dens)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~ pop, nrow = 1, scales = "free") +
  labs(title = "Population Distributions for Each Simulation Setting")
```

## Normal Min

```{r}
n <- 5
mu <- 10
sigma <- 2
# sample size
# population mean
# population standard deviation
generate_samp_min <- function(mu, sigma, n) {
   single_sample <- rnorm(n, mu, sigma)
  sample_min <- min(single_sample)
  return(sample_min)
}
## test function once:
generate_samp_min(mu = mu, sigma = sigma, n = n)
nsim <- 5000      # number of simulations
## code to map through the function.
## the \(i) syntax says to just repeat the generate_samp_mean function
## nsim times
mins <- map_dbl(1:nsim, \(i) generate_samp_min(mu = mu, sigma = sigma, n = n))
## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
norm_mins_df <- tibble(mins)
norm_mins_df

ggplot(data = norm_mins_df, aes(x = mins)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Mins",
       title = paste("Sampling Distribution of the \nSample Min when n =", n))
```

```{r}
norm_mins_df |>
  summarise(mean_samp_dist = mean(mins),
            var_samp_dist = var(mins),
            sd_samp_dist = sd(mins))
```

## Norm Max

```{r}
n <- 5
mu <- 10
sigma <- 2
# sample size
# population mean
# population standard deviation
generate_samp_max <- function(mu, sigma, n) {
   single_sample <- rnorm(n, mu, sigma)
  sample_max <- max(single_sample)
  return(sample_max)
}
## test function once:
generate_samp_max(mu = mu, sigma = sigma, n = n)
nsim <- 5000      # number of simulations
## code to map through the function.
## the \(i) syntax says to just repeat the generate_samp_mean function
## nsim times
maxs <- map_dbl(1:nsim, \(i) generate_samp_max(mu = mu, sigma = sigma, n = n))
## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
maxs_df <- tibble(maxs)


ggplot(data = maxs_df, aes(x = maxs)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Mins",
       title = paste("Sampling Distribution of the \nSample Max when n =", n))
```

```{r}
maxs_df |>
  summarise(mean_samp_dist = mean(maxs),
            var_samp_dist = var(maxs),
            sd_samp_dist = sd(maxs))
```

## Unif Min

```{r}
n <- 5
theta1 <- 7
theta2 <- 13
# sample size
# population mean
# population standard deviation
generate_unif_samp_min <- function(n, theta1, theta2) {
   single_sample <- runif(n, theta1, theta2)
  sample_min <- min(single_sample)
  return(sample_min)
}
## test function once:
generate_unif_samp_min(n = n, theta1 = theta1, theta2 = theta2)
nsim <- 5000      # number of simulations
## code to map through the function.
## the \(i) syntax says to just repeat the generate_samp_mean function
## nsim times
mins <- map_dbl(1:nsim, \(i) generate_unif_samp_min(n = n, theta1 = theta1, theta2 = theta2))
## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
unif_mins_df <- tibble(mins)
unif_mins_df

ggplot(data = unif_mins_df, aes(x = mins)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Mins",
       title = paste("Sampling Distribution of the \nSample Min when n =", n))
```

```{r}
unif_mins_df |>
  summarise(mean_samp_dist = mean(mins),
            var_samp_dist = var(mins),
            sd_samp_dist = sd(mins))
```

## Unif Max

```{r}
n <- 5
theta1 <- 7
theta2 <- 13
# sample size
# population mean
# population standard deviation
generate_unif_samp_max <- function(n, theta1, theta2) {
   single_sample <- runif(n, theta1, theta2)
  sample_max <- max(single_sample)
  return(sample_max)
}
## test function once:
generate_unif_samp_max(n = n, theta1 = theta1, theta2 = theta2)
nsim <- 5000      # number of simulations
## code to map through the function.
## the \(i) syntax says to just repeat the generate_samp_mean function
## nsim times
maxs <- map_dbl(1:nsim, \(i) generate_unif_samp_max(n = n, theta1 = theta1, theta2 = theta2))
## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
unif_max_df <- tibble(maxs)
unif_max_df

ggplot(data = unif_max_df, aes(x = maxs)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Maxs",
       title = paste("Sampling Distribution of the \nSample Max when n =", n))
```

```{r}
unif_max_df |>
  summarise(mean_samp_dist = mean(maxs),
            var_samp_dist = var(maxs),
            sd_samp_dist = sd(maxs))
```

## Exp Min

```{r}
n <- 5       # sample size
lambda <- 0.5
mu <- 1 / lambda   # population mean
sigma <- sqrt(1 / lambda ^ 2)  # population standard deviation
generate_exp_min <- function(lambda, n) {
  single_sample <- rexp(n, lambda)
  sample_min <- min(single_sample)
  return(sample_min)
}

nsim <- 5000      # number of simulations
mins <- map_dbl(1:nsim, \(i) generate_exp_min(lambda = lambda, n = n))

## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
exp_mins_df <- tibble(mins)

ggplot(data = exp_mins_df, aes(x = mins)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Mins",
       title = paste("Sampling Distribution of the \nSample Min when n =", n))
```

```{r}
exp_mins_df |>
  summarise(mean_samp_dist = mean(mins),
            var_samp_dist = var(mins),
            sd_samp_dist = sd(mins))
```

## Exp Max

```{r}
n <- 5       # sample size
lambda <- 0.5
mu <- 1 / lambda   # population mean
sigma <- sqrt(1 / lambda ^ 2)  # population standard deviation
generate_exp_max <- function(lambda, n) {
  single_sample <- rexp(n, lambda)
  sample_max <- max(single_sample)
  return(sample_max)
}

nsim <- 5000      # number of simulations
maxs <- map_dbl(1:nsim, \(i) generate_exp_max(lambda = lambda, n = n))

## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
exp_max_df <- tibble(maxs)

ggplot(data = exp_max_df, aes(x = maxs)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Maxs",
       title = paste("Sampling Distribution of the \nSample Max when n =", n))
```

```{r}
exp_max_df |>
  summarise(mean_samp_dist = mean(maxs),
            var_samp_dist = var(maxs),
            sd_samp_dist = sd(maxs))
```

## Beta Min

```{r}
n <- 5       # sample size
alpha <- 8
beta <- 2
mu <- alpha / (alpha + beta)   # population mean
sigma <- sqrt((alpha * beta) / ((alpha + beta)^2 * (alpha + beta + 1))) # population standard deviation

generate_beta_min <- function(n, alpha, beta) {
  single_sample <- rbeta(n, alpha, beta)
  sample_min <- min(single_sample)
  return(sample_min)
}

nsim <- 5000      # number of simulations
mins <- map_dbl(1:nsim, \(i) generate_beta_min(n = n, alpha = alpha, beta = beta))

## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
beta_mins_df <- tibble(mins)

ggplot(data = beta_mins_df, aes(x = mins)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Mins",
       title = paste("Sampling Distribution of the \nSample Min when n =", n))

```

```{r}
beta_mins_df |>
  summarise(mean_samp_dist = mean(mins),
            var_samp_dist = var(mins),
            sd_samp_dist = sd(mins))
```

## Beta Max

```{r}
n <- 5       # sample size
alpha <- 8
beta <- 2
mu <- alpha / (alpha + beta)   # population mean
sigma <- sqrt((alpha * beta) / ((alpha + beta)^2 * (alpha + beta + 1))) # population standard deviation

generate_beta_max <- function(n, alpha, beta) {
  single_sample <- rbeta(n, alpha, beta)
  sample_max <- max(single_sample)
  return(sample_max)
}

nsim <- 5000      # number of simulations
maxs <- map_dbl(1:nsim, \(i) generate_beta_max(n = n, alpha = alpha, beta = beta))

## print some of the 5000 means
## each number represents the sample mean from __one__ sample. means_df <- tibble(means)
beta_maxs_df <- tibble(maxs)

ggplot(data = beta_maxs_df, aes(x = maxs)) +
  geom_histogram(colour = "deeppink4", fill = "deeppink1", bins = 20) +
  theme_minimal() +
  labs(x = "Observed Sample Mins",
       title = paste("Sampling Distribution of the \nSample Maxs when n =", n))

```

```{r}
beta_maxs_df |>
  summarise(mean_samp_dist = mean(maxs),
            var_samp_dist = var(maxs),
            sd_samp_dist = sd(maxs))
```

|                      | $\text{N}(\mu = 10, \sigma^2 = 4)$ | $\text{Unif}(\theta_1 = 7, \theta_2 = 13)$ | $\text{Exp}(\lambda = 0.5)$ | $\text{Beta}(\alpha = 8, \beta = 2)$ |
|:------------:|:--------------:|:------------:|:------------:|:------------:|
| $\text{E}(Y_{min})$  |                7.70                |                    8.00                    |            0.389            |                0.646                 |
| $\text{E}(Y_{max})$  |                12.3                |                   12.00                    |            4.58             |                0.922                 |
|                      |                                    |                                            |                             |                                      |
| $\text{SE}(Y_{min})$ |                1.33                |                   0.838                    |            0.385            |                0.105                 |
| $\text{SE}(Y_{max})$ |                1.33                |                    0.87                    |            2.41             |                0.0452                |

: Table of Results {.striped .hover}

## Question 1

For the normal distribution with mu = 10 and sigma = 2, the distribution of the sample minimum and maximum when n = 5 have roughly the same Standard Error which makes sense becuase they both come from a normally distributed population.

For the Uniform distribution with theta one = 7 and theta 2 = 13, the distribution of the sample minimum and maximum when n = 5 have roughly the same Standard Error also. This is also apparent when you look at the distributions respective graphs, they have the same skewed shape but are basically mirrored versions of each other.

For the exponential distribution with lambda = 0.5, the distributions of the sample minimum and maximum when n = 5 have different Standard Errors. The Standard Error for the Sample Minimum distribution is much smaller than the Standard Error for the Sample Maximum distribution. This can be seen visually when comparing the histograms of the sample minimum and maximum distributions as the sample maximum looks like a more stretched out version of the sample minimum distribution.

I also think this is intuitive when looking at the graph of the population distribution. There's more density towards the lower values, so more of the minimums will be sampled close to and it will create a right-skewed distribution with a lot of density close to 0. The density of the maximum will be more spread out considering the shape of the population graph, which we also can see in the histogram od the sample maximum distribution. I expect this to hold for a similar population to the one we tested with.

For the beta distribution with alpha = 8 and beta = 2, the distributions of the sample minimum and maximum when n = 5 have different Standard Errors. This time it's the other way around and the Standard Error for the Sample Minimum is bigger than the Standard Error for the Sample Maximum Distribution.

When comparing the population graphs between the Beta and Exponential population I belive you can see why this is the result. The Beta graph has low density close to 0 and then it accelrates around 0.5 up to almost 1. This for me means that the sample minimum might be spread out while the sample maximum might be more concentrated which we also can see in their repsective graphs. I also expect this to hold for a similar population to the one we tested with.

## Question 2

## Exp min

```{r}
n <- 5
## CHANGE 0 and 3 to represent where you want your graph to start and end
## on the x-axis
x <- seq(0, 5, length.out = 1000)
## CHANGE to be the pdf you calculated. Note that, as of now, 
## this is not a proper density (it does not integrate to 1).
density_min <- 2.5 * exp(-2.5 * x)
  
## put into tibble and plot
samp_min_df <- tibble(x, density_min)
ggplot(data = samp_min_df, aes(x = x, y = density_min)) +
  geom_line() +
  theme_minimal()
```

## Exp Max

```{r}
n <- 5
## CHANGE 0 and 3 to represent where you want your graph to start and end
## on the x-axis
x <- seq(0, 20, length.out = 1000)
## CHANGE to be the pdf you calculated. Note that, as of now, 
## this is not a proper density (it does not integrate to 1).
density_max <- 2.5 * ((1 - exp(-0.5 * x))^4 * exp(-0.5 * x))
  
## put into tibble and plot
samp_max_df <- tibble(x, density_max)
ggplot(data = samp_max_df, aes(x = x, y = density_max)) +
  geom_line() +
  theme_minimal()
```

# Find Expectation by integrating both dnesity fucntions

```{r}
# Integrate the pdf of the sample maximum distribution * x from 0 to infinity 
ev_max <- integrate(function(x) {x * (2.5 * ((1 - exp(-0.5 * x))^4 * exp(-0.5 * x)))}
                    , lower = 0, upper = Inf)
ev_max

# Integrate the pdf of the sample minimum distribution * x from 0 to infinity 
ev_min <- integrate(function(x) {x * (2.5 * exp(-2.5 * x))}
                    ,lower = 0, upper = Inf)
ev_min

# Integrate the pdf of the sample maximum distribution * x^2 from 0 to infinity 
ev_squared_max <- integrate(function(x) {x^2 * (2.5 * ((1 - exp(-0.5 * x))^4 * exp(-0.5 * x)))}
                            ,lower = 0, upper = Inf)
ev_squared_max

# Integrate the pdf of the sample minimum distribution * x^2 from 0 to infinity 
ev_squared_min <- integrate(function(x) {x^2 * (2.5 * exp(-2.5 * x))}
                            ,lower = 0, upper = Inf)
ev_squared_min

# Use the results to calculate variance, take square root to get SE
se_max <- sqrt(26.70889 - 4.566667^2)
se_max

# Use the results to calculate variance, take square root to get SE
se_min <- sqrt(0.32 - 0.4^2)
se_min
```

|                      | $\text{Exp}(\lambda = 0.5)$ |     |
|:--------------------:|:---------------------------:|:---:|
| $\text{E}(Y_{min})$  |             0.4             |     |
| $\text{E}(Y_{max})$  |            4.567            |     |
|                      |                             |     |
| $\text{SE}(Y_{min})$ |             0.4             |     |
| $\text{SE}(Y_{max})$ |            2.419            |     |

: Table of Results {.striped .hover}

I think my results are fairly close to the simulated answers. The expectation of the sample minimum distribution was off by 0.11, the expectation of the sample maximum distribution was off by 0.013. The Standard Error of the minimum distribution was off by 0.015 and the standard error of the maximum distribution was the same except for some rounding. I think this is a cool result and goes to show that simulation is a useful tool to find sampling distributions of many different statistics from different populations.
