---
title: "Mini Project 3: Confidence Interval Assumptions"
---

```{r}
library(tidyverse)
```

n small : 10 n medium : 25 n large: 250

p1 = 0.47 p2 = 0.9

## Large N, p = 0.9

```{r}
n <- 250   # sample size
p <- 0.9  # population proportion
  
generate_samp_prop <- function(n, p) {
  x <- rbinom(1, n, p) # randomly generate number of successes for the sample
  
  ## number of successes divided by sample size
  phat <- x / n
  
  # 90% CI
  lb <- phat - 1.645 * sqrt(phat * (1- phat) / n)
  ub <- phat + 1.645 * sqrt(phat * (1- phat) / n)
  
  # return df with sample proportion, lower and upper bounds of CI for that sample
  prop_df <- tibble(phat, lb, ub)
  return(prop_df)
}

generate_samp_prop(n = 250, p = 0.9)

# how many CIs do we want
nsim <- 5000

prop_ci_df <- map(1:nsim, 
    \(i) generate_samp_prop(n = 250, p = 0.9)) |> bind_rows()

prop_ci_df
```

Average interval width:

```{r}
mean_w_large_p2 <- mean(prop_ci_df$ub - prop_ci_df$lb)
mean_w_large_p2
# True if phat between the bounds
# False otherwise
# mean of binary variable is the proprtion true values
cov_rate_large_p2 <-  mean(p > prop_ci_df$lb & p < prop_ci_df$ub)
cov_rate_large_p2
```

Check Large Sample Assumption

```{r}
n*p
n*(1 - p)
```

Large sample assumption holds in this case because both are greater than 10.

## Large N, p = 0.47

```{r}
n <- 250   # sample size
p <- 0.47  # population proportion
  
generate_samp_prop <- function(n, p) {
  x <- rbinom(1, n, p) # randomly generate number of successes for the sample
  
  ## number of successes divided by sample size
  phat <- x / n
  
  # 90% CI
  lb <- phat - 1.645 * sqrt(phat * (1- phat) / n)
  ub <- phat + 1.645 * sqrt(phat * (1- phat) / n)
  
  # return df with sample proportion, lower and upper bounds of CI for that sample
  prop_df <- tibble(phat, lb, ub)
  return(prop_df)
}

generate_samp_prop(n = 250, p = 0.47)

# how many CIs do we want
nsim <- 5000

prop_ci_df <- map(1:nsim, 
    \(i) generate_samp_prop(n = 250, p = 0.47)) |> bind_rows()

prop_ci_df
```

Average interval width:

```{r}
mean_w_large_p1 <- mean(prop_ci_df$ub - prop_ci_df$lb)
mean_w_large_p1
# True if phat between the bounds
# False otherwise
# mean of binary variable is the proprtion true values
cov_rate_large_p1 <- mean(p > prop_ci_df$lb & p < prop_ci_df$ub)
cov_rate_large_p1
```

Check Large Sample Assumption

```{r}
n*p
n*(1 - p)
```

Large sample assumption holds in this case because both are greater than 10.

## Medium N, p = 0.9

```{r}
n <- 25   # sample size
p <- 0.9  # population proportion
  
generate_samp_prop <- function(n, p) {
  x <- rbinom(1, n, p) # randomly generate number of successes for the sample
  
  ## number of successes divided by sample size
  phat <- x / n
  
  # 90% CI
  lb <- phat - 1.645 * sqrt(phat * (1- phat) / n)
  ub <- phat + 1.645 * sqrt(phat * (1- phat) / n)
  
  # return df with sample proportion, lower and upper bounds of CI for that sample
  prop_df <- tibble(phat, lb, ub)
  return(prop_df)
}

generate_samp_prop(n = 25, p = 0.9)

# how many CIs do we want
nsim <- 5000

prop_ci_df <- map(1:nsim, 
    \(i) generate_samp_prop(n = 25, p = 0.9)) |> bind_rows()

prop_ci_df
```

Average interval width:

```{r}
mean_w_medium_p2 <- mean(prop_ci_df$ub - prop_ci_df$lb)
mean_w_medium_p2
# True if phat between the bounds
# False otherwise
# mean of binary variable is the proprtion true values
cov_rate_medium_p2 <-  mean(p > prop_ci_df$lb & p < prop_ci_df$ub)
cov_rate_medium_p2
```

Check Large Sample Assumption

```{r}
n*p
n*(1 - p)
```

Large sample assumption does not hold in this case because n \*(1-p) is less than 10. Therefore, the asymptotic confidence interval might not be accurate. This is because the sample is highly skewed because p = 0.9 and n is relatively small.

## Medium N, p = 0.47

```{r}
n <- 25   # sample size
p <- 0.47  # population proportion
  
generate_samp_prop <- function(n, p) {
  x <- rbinom(1, n, p) # randomly generate number of successes for the sample
  
  ## number of successes divided by sample size
  phat <- x / n
  
  # 90% CI
  lb <- phat - 1.645 * sqrt(phat * (1- phat) / n)
  ub <- phat + 1.645 * sqrt(phat * (1- phat) / n)
  
  # return df with sample proportion, lower and upper bounds of CI for that sample
  prop_df <- tibble(phat, lb, ub)
  return(prop_df)
}

generate_samp_prop(n = 25, p = 0.47)

# how many CIs do we want
nsim <- 5000

prop_ci_df <- map(1:nsim, 
    \(i) generate_samp_prop(n = 25, p = 0.47)) |> bind_rows()

prop_ci_df
```

Average interval width:

```{r}
mean_w_medium_p1 <- mean(prop_ci_df$ub - prop_ci_df$lb)
mean_w_medium_p1
# True if phat between the bounds
# False otherwise
# mean of binary variable is the proprtion true values
cov_rate_medium_p1 <- mean(p > prop_ci_df$lb & p < prop_ci_df$ub)
cov_rate_medium_p1
```

Check Large Sample Assumption

```{r}
n*p
n*(1 - p)
```

The large sample assumption holds because both values are greater than 10. l \## Small n, p = 0.9

```{r}
n <- 10   # sample size
p <- 0.9  # population proportion
  
generate_samp_prop <- function(n, p) {
  x <- rbinom(1, n, p) # randomly generate number of successes for the sample
  
  ## number of successes divided by sample size
  phat <- x / n
  
  # 90% CI
  lb <- phat - 1.645 * sqrt(phat * (1- phat) / n)
  ub <- phat + 1.645 * sqrt(phat * (1- phat) / n)
  
  # return df with sample proportion, lower and upper bounds of CI for that sample
  prop_df <- tibble(phat, lb, ub)
  return(prop_df)
}

generate_samp_prop(n = 10, p = 0.9)

# how many CIs do we want
nsim <- 5000

prop_ci_df <- map(1:nsim, 
    \(i) generate_samp_prop(n = 10, p = 0.9)) |> bind_rows()

prop_ci_df
```

Average interval width:

```{r}
mean_w_small_p2 <- mean(prop_ci_df$ub - prop_ci_df$lb)
mean_w_small_p2
# True if phat between the bounds
# False otherwise
# mean of binary variable is the proprtion true values
cov_rate_small_p2 <- mean(p > prop_ci_df$lb & p < prop_ci_df$ub)
cov_rate_small_p2
```

Check Large Sample Assumption

```{r}
n*p
n*(1 - p)
```

The large sample assumption does not hold. In this case, n is small and the distribution of p is not normally distributed so I think it fails because of both.

## Small n, p = 0.47

```{r}
n <- 10   # sample size
p <- 0.47  # population proportion
  
generate_samp_prop <- function(n, p) {
  x <- rbinom(1, n, p) # randomly generate number of successes for the sample
  
  ## number of successes divided by sample size
  phat <- x / n
  
  # 90% CI
  lb <- phat - 1.645 * sqrt(phat * (1- phat) / n)
  ub <- phat + 1.645 * sqrt(phat * (1- phat) / n)
  
  # return df with sample proportion, lower and upper bounds of CI for that sample
  prop_df <- tibble(phat, lb, ub)
  return(prop_df)
}

generate_samp_prop(n = 10, p = 0.47)

# how many CIs do we want
nsim <- 5000

prop_ci_df <- map(1:nsim, 
    \(i) generate_samp_prop(n = 10, p = 0.47)) |> bind_rows()

prop_ci_df
```

Average interval width:

```{r}
mean_w_small_p1 <- mean(prop_ci_df$ub - prop_ci_df$lb)
mean_w_small_p1
# True if phat between the bounds
# False otherwise
# mean of binary variable is the proprtion true values
cov_rate_small_p1 <- mean(p > prop_ci_df$lb & p < prop_ci_df$ub)
cov_rate_small_p1
```

Check Large Sample Assumption

```{r}
n*p
n*(1 - p)
```

The large sample assumption does not hold in this case. Even if p is supposed to relatively normally distributed, with such a small n the large sample assumption will not hold.

## Table

|            |               | $n = 250$ | $n = 25$ | $n = 10$ |
|:----------:|:-------------:|:---------:|:--------:|:--------:|
| $p = 0.9$  | Coverage Rate |  0.8822   |  0.9182  |  0.6392  |
| $p = 0.47$ | Coverage Rate |  0.8992   |  0.8932  |  0.8818  |
|            |               |           |          |          |
| $p = 0.9$  | Average Width |   0.062   |  0.1804  |  0.2358  |
| $p = 0.47$ | Average Width |  0.1037   |  0.3217  |  0.4904  |

: Table of Results {.striped .hover}

## Findings

This project showed me how the theoretical assumptions of confidence intervals work and how they connect to real-life scenarios. It is clear that with a very big n, such as n = 250, the asymptotic confidence interval holds. Intuitively I was surprised that the coverage rate was higher for n = 25 than for n = 250 when p = 0.9. However, I realized that this was because the average width of the interval got wider as n decreased, since the margin of error formula includes division by n. A wider interval has a higher chance of capturing p, leading to a higher coverage rate.

I also thought it was interesting that the average width of the interval was smaller for p = 0.9 than for p = 0.47. This is because the standard error is bigger when p = 0.47 since 0.47(1-0.47) \> 0.9(1-0.9). However, this does not always translate to a higher coverage rate for p = 0.47 which we can see for n = 25.

Another interesting takeaway for me was that when the large sample assumtpion is broken but the data is approximately normal, it seems like the asymptotic confidence interval is still relatively correct. For example, the large sample assumption is broken for n = 10 and p = 0.47 but the coverage rate is still close to 0.9. Compared to when n = 10 and p = 0.9, since binom with p = 0.9 is skeewed, the small n is not enough to make up for this skewness and the coverage rate is only 0.63. Overall, this project helped me understand not just why the assumptions matter, but also how breaking them affects the accuracy of confidence intervals in different ways.
